Making Sequential Programs Deterministic

### 2019-03-22 design.tex
design subsection: dataflow analysis

\subsection{Dataflow analysis and type refinement}
The checker framework performs automatic type inference for local types within method bodies.
It also performs dataflow analysis and refines the types of these locals, thereby reducing the annotation burden 
on the programmer.

\begin{itemize}
	\item Example code with explicit annotations on local variable, showing flow of values into this local variable
	\item Explain that CF refines the type of this local variable; inbuild dataflow analysis
	\item Constrast this with a type on a method argument that won't be type refined.
	\item CF allows type system developers to define additional refinement rules.
	\item Determinism checker performs type refinement for the receivers of sort() methods in Collection classes.
	\item Example client code where this type refinement results in precise results
\end{itemize}

>>>>

\subsection{Dataflow analysis and type refinement}
The checker framework performs automatic type inference for local types within method bodies.
It also performs dataflow analysis and refines the types of these locals, thereby reducing the annotation burden 
on the programmer. Consider the example below:
\begin{verbatim}
void foo(@Det int x) {
    @NonDet int y = x;    // Type of y gets refined to @Det.
    @Det int z = y;       //  No error here.
}
\end{verbatim}
The checker framework performs dataflow analysis within method bodies and refines the types of local variables.
In the example above, even though the local variable \codeid{y} was declared to be \codeid{@NonDet}, it gets
type refined to \codeid{@Det} after the first assignment statement. As the result, the assignment of \codeid{y}
to a \codeid{@Det} variable \codeid{z} is sound. In the follwing example,
\begin{verbatim}
void foo(@NonDet Object x, @Det Object y) {
    x = y;                // No type refinement for x.
    @Det Object z = x;    // Error.
}
\end{verbatim}
the checker reports a warning at the second assignment statement of \codeid{foo}. Type refinement only applies to locals. Refining non-locals can result in side effects making the analysis unsound.

The determinism checker has additional type refinement rules for the sorting methods in collections. 
The checker type refines a receiver of type \codeid{@OrderNonDet List} to a \codeid{@Det List} if all the
type argument of this \codeid{List} have the type annotation \codeid{@Det}. Similar rules apply for \codeid{Arrays.sort()} and
\codeid{Collections.sort()}.
Under this if \codeid{sort()} is called on a receiver of type \codeid{@OrderNonDet List<@Det Set<@Det String>>}, it will
be type refined to \codeid{@Det List<@Det Set<@Det String>>}. However a receiver of type \codeid{@OrderNonDet List<@OrderNonDet Set<@Det String>>} will not be type refined.
